---
title: "Results Aging"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: 
      collapsed: false
    includes:
      in_header: ../include/in_header.html
      before_body: ../include/before_body.html
      after_body: ../include/after_body.html
runtime: shiny_prerendered
---
<!-- # (c)  Juozas GordeviÄius -->

```{r, include=FALSE}
require(broom)
require(knitr)
require(RobustRankAggreg)
require(cowplot)
source("../code/common.R")
source("../code/pformat.R")
source("../code/genomicEnrichment.R")
annotations <- cache(foo=loadHumanAnnotations,
                     fname = "annotations.RDS")
pathways <- readRDS("../Discover_Pathways/humanPathways.RDS")
```

# Appendix, padlocks

```{r}
apax <- fread("../Appendix_AgeAcceleration_Padlock/www/m6_Appendix_AgeAcceleration_Padlock.csv")
apad <- readRDS("../Appendix_AgeAcceleration_Padlock/www/m6_methylation.RDS")
```

```{r, include=TRUE, results='asis'}
n <- nrow(apax)
glue("- Number of loci {n}")
t <- apad$key[, table(Diagnosis)]
glue("- Number of PD cases: {t[2]} and controls: {t[1]}")
n <- apax[, sum(p.adjust(P.DiagnosisControl.Age, "fdr") < 0.05, na.rm=TRUE)]
glue("- Number of significant loci {n}")
n <- apax[p.adjust(P.DiagnosisControl.Age, "fdr") < 0.05, length(unique(Gene))]
glue("- Number of genes affected {n}")
r <- apad$key[Diagnosis == "Control", range(Age) %>% paste(collapse=":")]
glue("- Age range of healthy controls {r}")
```


Dominant direction of methylation

```{r, include=TRUE}
t <-
	apax[, list(
						Significant = ifelse(p.adjust(P.DiagnosisControl.Age, "fdr") < 0.05, "Significant", "NS"), 
					 	Direction = ifelse(C.DiagnosisControl.Age < 0, "Hypo-M", "Hyper-M")
					 )] %>%
	.[, table(Significant, Direction)] %>%
	.[c("NS", "Significant"), c("Hypo-M", "Hyper-M")]

# Show the table
knitr::kable(t) %>%
kable_styling(bootstrap_options = c("hover"),
							full_width = FALSE)

# Show fisher test result
t %>% fisher.test %>% 
tidy() %>%
kable() %>%
kable_styling(bootstrap_options = c("hover"))
```


```{r}
# touch()
foo <- function(analysis, prefix, annotations, t) {
  fits <- fread(glue("../{analysis}/www/{prefix}_{analysis}.csv"))
  ccol <- glue("C.DiagnosisControl.Age")
  pcol <- glue("P.DiagnosisControl.Age")
  fits <- fits[, list(ID, Chr, SNP, Str, 
                  adj.P.Val = p.adjust(get(pcol), "fdr"),
                  logFC = get(ccol)
                  )]
  dt1 <- getAnnotatrEnrichments(fits, annotations, qThresh = t)
  dt2 <- getAppendixChipEnrichments(fits, qThresh = t)
  dt3 <- getEpiRoadmapEnrichment(fits, qThresh = t)
  dt4 <- getPsychEncodeEnrichments(fits, qThresh = t)
  return(list(dt1, dt2, dt3, dt4))
}

# touch()
analysis <- "Appendix_AgeAcceleration_Padlock"
prefix <- "m6"
t <- 0.05

dtlist <- cache(
  foo = foo, fname = glue("enrichment_{analysis}.RDS"), verbose = TRUE,
  analysis = analysis, prefix= prefix, annotations = annotations,
  t = t
  )
```


General annotation of genomic elements:

```{r, include=TRUE}
dtlist[[1]] %>%
.[Type == "All"] %>%
.[, `P < 0.05` := texPstars(`P < 0.05`)] %>%
dplyr::select(estimate=OR, 
       p.value=P, 
       conf.low=Lo, 
       conf.high=Hi, 
       category=Category, 
       p_stars = `P < 0.05`) %>%
kable() %>%
kable_styling(bootstrap_options = c("hover"))
```

Genomic elements obtained from ChIP-seq:

```{r, include=TRUE}
dtlist[[2]] %>%
.[, `P < 0.05` := texPstars(`P < 0.05`)] %>%
dplyr::select(estimate=OR, 
       p.value=P, 
       conf.low=Lo, 
       conf.high=Hi, 
       category=Category, 
       type=Type,
       p_stars = `P < 0.05`) %>%
kable() %>%
kable_styling(bootstrap_options = c("hover"))
```

```{r}
myplots <- plotAnnotatrEnrichment(dtlist[[1]])
myplots[[1]] <- myplots[[1]] + facet_grid(Type~.)
myplots[[2]] <- myplots[[2]] + facet_grid(Type~.)

myplots[[3]] <- plotAppendixChipEnrichment(dtlist[[2]], wrap = "vertical")
myplots[[4]] <- plotEpiRoadmapEnrichment(  dtlist[[3]], wrap = "vertical")
```

```{r, include=TRUE, fig.cap="**Figure S7. Genomic elements enriched with differentially methylated cytosines in the healthy aging appendix.**"}
ggarrange(myplots[[1]], myplots[[2]], myplots[[3]],
  ncol = 3, nrow = 1, labels = LETTERS[1:3],
  widths = c(4, 7, 3),
  common.legend = TRUE)
```


```{r}
url <- "../Appendix_AgeAcceleration_Padlock/www/m6_Appendix_AgeAcceleration_Padlock.csv"
stopifnot(file.exists(url))
```



- **[Additional file 11](`r url`) DNA methylation changes in the aging appendix of healthy individuals and PD patients.**


# Appendix, RNA-seq

```{r}
apar <- fread("../Appendix_AgeAcceleration_RNAseq/www/r1_Appendix_AgeAcceleration_RNAseq.csv")
apard <- readRDS("../Appendix_AgeAcceleration_RNAseq/www/r1_rnaseq.RDS")
```

```{r, include=TRUE, results='asis'}
n <- nrow(apar)
glue("- Number of transcripts {n}")
t <- apard$key[, table(Diagnosis)]
glue("- Number of PD cases: {t[2]} and controls: {t[1]}")
n <- apar[, sum(p.adjust(P.DiagnosisControl.Age, "fdr") < 0.05, na.rm=TRUE)]
glue("- Number of significant transcripts {n}")
```


Dominant direction of transcription change:

```{r, include=TRUE}
t <-
	apar[, list(
						Significant = ifelse(p.adjust(P.DiagnosisControl.Age, "fdr") < 0.05, "Significant", "NS"), 
					 	Direction = ifelse(C.DiagnosisControl.Age < 0, "Down", "Up")
					 )] %>%
	.[, table(Significant, Direction)] %>%
	.[c("NS", "Significant"), c("Down", "Up")]

# Show the table
knitr::kable(t) %>%
kable_styling(bootstrap_options = c("hover"),
							full_width = FALSE)

# Show fisher test result
t %>% fisher.test %>% 
broom::tidy() %>%
knitr::kable() %>%
kable_styling(bootstrap_options = c("hover"))
```

# Prefrontal cortex neurons, padlock

```{r}
pfcax <- fread("../Brain_AgeAcceleration_Padlock/www/m6_Brain_AgeAcceleration_Padlock.csv")
pfcad <- readRDS("../Brain_AgeAcceleration_Padlock/www/m6_methylation.RDS")
```

```{r, include=TRUE, results='asis'}
n <- nrow(pfcax[Type == "CG"])
glue("- Number of CpG loci {n}")
n <- nrow(pfcax[Type != "CG"])
glue("- Number of CpH loci {n}")
t <- pfcad$key[, table(Group)]
glue("- Number of PD cases: {t[2]} and controls: {t[1]}")
n <- pfcax[, sum(p.adjust(P.GroupCTRL.Age, "fdr") < 0.05, na.rm=TRUE)]
glue("- Number of significant loci {n}")
n <- pfcax[p.adjust(P.GroupCTRL.Age, "fdr") < 0.05, length(unique(Gene))]
glue("- Number of genes affected {n}")
```

Dominant direction of methylation

```{r, include=TRUE}
t <-
	pfcax[, list(
						Significant = ifelse(p.adjust(P.GroupCTRL.Age, "fdr") < 0.05, "Significant", "NS"), 
					 	Direction = ifelse(C.GroupCTRL.Age < 0, "Hypo-M", "Hyper-M")
					 )] %>%
	.[, table(Significant, Direction)] %>%
	.[c("NS", "Significant"), c("Hypo-M", "Hyper-M")]

# Show the table
knitr::kable(t) %>%
kable_styling(bootstrap_options = c("hover"),
							full_width = FALSE)

# Show fisher test result
t %>% fisher.test %>% 
tidy() %>%
kable() %>%
kable_styling(bootstrap_options = c("hover"))
```

```{r}
url <- "../Brain_AgeAcceleration_Padlock/www/m6_Brain_AgeAcceleration_Padlock.csv"
stopifnot(file.exists(url))
```

- **[Additional file 12](`r url`) DNA methylation changes in aging prefrontal cortex neurons of healthy individuals and PD patients.**


# Enrichment of ALP pathways in the epigenome

```{r}
pd <- foreach(p = pathways, .combine = rbind) %do% {
  genes <- p$genes
  colP  <- "P.DiagnosisControl.Age"
  colFC <- "C.DiagnosisControl.Age"
  dt <- apax[, 
          list(
            Q = get(colP) %>% p.adjust("fdr"),
            FC= get(colFC),
            Gene)]
  res1 <- 
        dt[, table(Q < 0.05, Gene %in% genes)] %>%
          fisher.test %>%
          broom::tidy() %>%
          setDT %>%
          .[, Term := p$term] %>%
          .[, Name := p$name] %>%
          .[, Type := "Appendix"] %>%
          .[, Dir  := "All"]
  res1up <- 
        dt[, table(Q < 0.05 & sign(FC) > 0, Gene %in% genes)] %>%
          fisher.test %>%
          broom::tidy() %>%
          setDT %>%
          .[, Term := p$term] %>%
          .[, Name := p$name] %>%
          .[, Type := "Appendix"] %>%
          .[, Dir  := "Hyper-"]
  res1down <- 
        dt[, table(Q < 0.05 & sign(FC) < 0, Gene %in% genes)] %>%
          fisher.test %>%
          broom::tidy() %>%
          setDT %>%
          .[, Term := p$term] %>%
          .[, Name := p$name] %>%
          .[, Type := "Appendix"] %>%
          .[, Dir  := "Hypo-"]

  colP  <- "P.GroupCTRL.Age"
  colFC <- "C.GroupCTRL.Age"
  dt <- pfcax[, 
          list(
            Q = get(colP) %>% p.adjust("fdr"),
            FC= get(colFC),
            Gene)]
  res2 <- 
        dt[, table( Q < 0.05, Gene %in% genes)] %>% 
          fisher.test %>%
          broom::tidy() %>%
          setDT %>%
          .[, Term := p$term] %>%
          .[, Name := p$name] %>%
          .[, Type := "PFC"] %>%
          .[, Dir  := "All"]
  res2up <- 
        dt[, table( Q < 0.05 & sign(FC) > 0, Gene %in% genes)] %>% 
          fisher.test %>%
          broom::tidy() %>%
          setDT %>%
          .[, Term := p$term] %>%
          .[, Name := p$name] %>%
          .[, Type := "PFC"] %>%
          .[, Dir  := "Hyper-"]

  res2down <- 
        dt[, table( Q < 0.05 & sign(FC) < 0, Gene %in% genes)] %>% 
          fisher.test %>%
          broom::tidy() %>%
          setDT %>%
          .[, Term := p$term] %>%
          .[, Name := p$name] %>%
          .[, Type := "PFC"] %>%
          .[, Dir  := "Hypo-"]

  rbindlist(list(res1, res1up, res1down, res2, res2up, res2down))
}
if (!dir.exists("www")) dir.create("www")
fwrite(pd, file = "www/enrichment_Pathways_DMCs.csv")

```

```{r, include=TRUE}
pd %>%
.[, `P < 0.05` := gtools::stars.pval(p.value)] %>% 
.[order(Type, Dir), ] %>%
.[, list(estimate, p.value, conf.low, conf.high, Type,
				 term = sprintf("%s (%s)", Name, Term),
				 direction = Dir, 
				 p_stars = texPstars(`P < 0.05`))] %>%
kable() %>%
kable_styling(bootstrap_options = c("hover"))
```

```{r, include=TRUE, fig.cap="**Figure S8. ALP pathway changes in DNA methylation in the healthy aging appendix and prefrontal cortex neurons.**"}
ggplot(pd, aes(Name, log(estimate), color = p.value < 0.05)) +
  geom_point(size = 4) + 
  geom_errorbar(aes(ymin = log(conf.low), ymax = log(conf.high)), width = 0.25, size = 0.25, color = "black") +
  geom_hline(yintercept = 0) + 
  ylab("OR, log") +
  scale_color_brewer("P < 0.05", palette="Dark2") +
  facet_grid(Type ~ Dir) + 
  theme_bw(base_size = 14) + 
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 55, hjust = 1),
        legend.position = "top") 
```


Correlation of odds ratios in PFC and appendix

```{r, include=TRUE}
pd[Dir == "Hyper-", list(Name, Type, estimate)] %>%
dcast(Name ~ Type, value.var = 'estimate') %>%
.[, cor.test(Appendix, PFC)] %>%
tidy() %>%
kable %>% 
kable_styling(bootstrap_options="hover")
```

# Aging hallmarks

```{r}
computeORs <- function(dt) {
	require(doParallel)
	registerDoParallel(cores = parallel::detectCores())
  dt <- 
    foreach (gene = unique(dt$Gene), .combine = rbind) %dopar% {
      require(data.table)
      tryCatch({
        t <- dt[, table(Gene == gene, Significant == TRUE)]
        f <- fisher.test(t, alternative = "greater")
        data.table(OR = f$estimate, P = f$p.value, Gene = gene)    
      }, error = function(e) {
        message(e)
        data.table(OR = NA, P = NA, Gene = gene)
      })
    }
  dt
}

epfc <- cache(foo=computeORs, fname="epfc.RDS",
              dt = pfcax[, list(Gene, Significant = p.adjust(P.GroupCTRL.Age, "fdr") < 0.05)])
eapx  <- cache(foo=computeORs, fname="eapx.RDS",
               dt = apax[, list(Gene, Significant = p.adjust(P.DiagnosisControl.Age, "fdr") < 0.05)])
hallmarks <- aggregateRanks(
  list(
    epfc[order(P), Gene],
    eapx[order(P), Gene]
  )
)  
setDT(hallmarks)
setnames(hallmarks, "Name", "Gene")
hallmarks <- 
merge(hallmarks, epfc[, list(Gene, PFC.OR=OR, PFC.P=P)], by = "Gene") %>% 
		merge(., eapx[, list(Gene, APX.OR=OR, APX.P=P)], by = "Gene")

if (!dir.exists("www")) dir.create("www")
write.table(hallmarks[order(Score)], file = "www/Aging_Hallmarks.csv", sep = ",", row.names = FALSE)

```

```{r, include=TRUE}
hallmarks[order(Score)][1:15] %>%
kable %>%
kable_styling(bootstrap_options="hover")
```

```{r}
url <- "./www/Aging_Hallmarks.csv"
stopifnot(file.exists(url))
```

- **[Additional file 13](`r url`) ALP genes consistently epigenetically altered with aging in the healthy appendix and healthy prefrontal cortex neurons.**


# Aging in PD and control individuals

```{r}
computeTrends <- function(X, fit, colP, colFC, age, group) {
  n <- nrow(X)
  ids <- fit[ get(colFC) > 0][order(get(colP)), ID][1:n]
  ids <- ids[ ids %in% rownames(X) ]
  w <- fit[ID %in% ids, get(colFC)]
  w <- w / sum(w)
  trend <- (X[ids,] * w) %>% colSums(na.rm=TRUE)
  pd <- data.table(Dir = "Hyper-M", Age = age, 
    Trend = trend, Group = group)

  ids <- fit[ get(colFC) < 0][order(get(colP)), ID][1:n]
  ids <- ids[ ids %in% rownames(X) ]
  w <- fit[ID %in% ids, get(colFC)]
  w <- w / sum(w)
  trend <- (X[ids,] * w) %>% colSums(na.rm=TRUE)
  pd <- rbind(pd, data.table(Dir = "Hypo-M", Age = age, 
    Trend = trend, Group = group))
  pd
}

# Appendix
apxdata <- readRDS("../Appendix_AgeAcceleration_Padlock/www/m6_methylation.RDS")
path <- "../Appendix_AgeAcceleration_Padlock/www/m6_Appendix_AgeAcceleration_Padlock.csv"
apxfit <- fread(path)

# restrict the data here
agerange <- apxdata$modelData[Diagnosis == "PD/LBD", range(Age)]
sidx <- apxdata$modelData[, Age >= agerange[1] & Age <= agerange[2] ]
apxdata <- shrinkList(apxdata, sidx, nrow(apxdata$modelData))

X <- lm( t(apxdata$methylation %>% na.omit) ~ ., 
  data = apxdata$modelData %>% select(-Age))$residuals %>% t
pd <- computeTrends(X, apxfit,
                    colP  = "P.DiagnosisControl.Age",
                    colFC = "C.DiagnosisControl.Age",
                    age   = apxdata$modelData$Age,
                    group = apxdata$modelData$Diagnosis) %>%
  .[, Type := "Appendix"] %>% 
  .[, Group := factor(Group, levels = c("Control", "PD/LBD"), labels = c("Control", "PD"))]


# PFC 
pfcdata <- readRDS("../Brain_AgeAcceleration_Padlock/www/m6_methylation.RDS")
path <- "../Brain_AgeAcceleration_Padlock/www/m6_Brain_AgeAcceleration_Padlock.csv"
pfcfit <- fread(path)

# restrict the data
agerange <- pfcdata$modelData[Group == "PD", range(Age)]
sidx <- pfcdata$modelData[, Age >= agerange[1] & Age <= agerange[2] ]
pfcdata <- shrinkList(pfcdata, sidx, nrow(pfcdata$modelData))



X <- lm( t(pfcdata$methylation %>% na.omit) ~ ., 
  data = pfcdata$modelData %>% select(-Age))$residuals %>% t
pd2 <- computeTrends(X, pfcfit,
                    colP  = "P.GroupCTRL.Age",
                    colFC = "C.GroupCTRL.Age",
                    age   = pfcdata$modelData$Age,
                    group = pfcdata$modelData$Group) %>%
  .[, Type := "PFC"] %>%
  .[, Group := factor(Group, levels = c("CTRL", "PD"), labels = c("Control", "PD"))]


pd <- rbind(pd, pd2)
textLayer <-
            pd[, cor.test(Age, Trend) %>% tidy, list(Group, Type, Dir)] %>%
            select(Group, Type, Dir, estimate, p.value) %>%
            setDT() %>%
            .[, label := 
                  sprintf("r = $%s$, p %s %s", 
                          format(estimate, digits=1), 
                          ifelse(p.value == 0, "<", "="),
                          texP(p.value)
                         )
            ] %>% 
            .[, Age := 20] %>% 
            .[, Trend := ifelse(Group == "Control", 0.9, 0.7)]
apxdata$modelData[, table(Diagnosis)]
pfcdata$modelData[, table(Group)]
```


```{r, include=TRUE, fig.cap="**Figure S9. Epigenetic aging of ALP genes in the appendix and prefrontal cortex neurons of age-matched healthy controls and PD cases.** "}
pd %>%
ggplot(., aes(Age, Trend, color = Group)) + 
  geom_point(size = 0.5, alpha = 0.3, aes(shape = Group, fill = Group)) + 
  geom_smooth(method = 'lm', size = 0.5, se = FALSE) + 
  geom_text(data = textLayer, aes(label = TeX(label, output="character")), 
    parse = TRUE, show.legend = FALSE,
    hjust=0) +
  ylab("Mean normalized methylation") + 
  facet_grid(Type ~ Dir, scales = "free_x") + 
  scale_shape_manual(values = c(16, 17)) +
  scale_color_manual(values = c("black", "#e31a1c")) +
  scale_fill_manual(values = c("black", "#e31a1c")) +
  scale_y_continuous(limits = c(-0.8, 1)) + 
  theme_bw(base_size = 14)
```

```{r}

# Aging vs PD effects in appendix RNAseq
alpGenes <- apax[, unique(Gene)]
path <- "../Appendix_PDvsControls_RNAseq/www/r1_Appendix_PDvsControls_RNAseq.csv"
apxpdfit <- fread(path)
apxpdfit <- apxpdfit[SYMBOL %in% alpGenes]

path <- "../Appendix_AgeAcceleration_RNAseq/www/r1_Appendix_AgeAcceleration_RNAseq.csv"
apxafit <- fread(path)
apxafit <- apxafit[SYMBOL %in% alpGenes]

pd <- merge(
            apxafit[, list(
                        SYMBOL, 
                        C = C.DiagnosisControl.Age, 
                        P=P.DiagnosisControl.Age, 
                        Q = p.adjust(P.DiagnosisControl.Age, "fdr"))],
            apxpdfit[, list(SYMBOL, logFC, P.Value, adj.P.Val)],
            by = "SYMBOL"
          ) %>%
  .[P < 0.05 & P.Value < 0.05, ] %>%
  .[, list(
          AgingEffect = abs(C), 
          PDEffect = abs(logFC),
          Type = "Appendix"
          )
  ]


pA <- 
  pd %>%
  ggplot(., aes(AgingEffect, PDEffect)) + 
    geom_point(shape = 1, color = "grey") +
    geom_smooth(method = "lm", color = "black") + 
    stat_cor(method = "pearson") +
    scale_x_continuous(
      labels = scales::number_format(accuracy = 0.01)) + 
    xlab("Absolute aging effect") + 
    ylab("Absolute PD effect") +
    theme_bw(base_size = 14)



# Normalized aging trend
computeTrends <- function(X, fit, colP, colFC, age, group) {
  n <- nrow(X)
  ids <- fit[ get(colFC) > 0][order(get(colP)), ENSEMBL][1:n]
  ids <- ids[ ids %in% rownames(X) ]
  w <- fit[ENSEMBL %in% ids, get(colFC)]
  w <- w / sum(w)
  trend <- (X[ids,] * w) %>% colSums(na.rm=TRUE)
  pd <- data.table(Dir = "Up-regulation", Age = age, 
    Trend = trend, Group = group)

  ids <- fit[ get(colFC) < 0][order(get(colP)), ENSEMBL][1:n]
  ids <- ids[ ids %in% rownames(X) ]
  w <- fit[ENSEMBL %in% ids, get(colFC)]
  w <- w / sum(w)
  trend <- (X[ids,] * w) %>% colSums(na.rm=TRUE)
  pd <- rbind(pd, data.table(Dir = "Down-regulation", Age = age, 
    Trend = trend, Group = group))
  pd
}

# Appendix

apxdata <- readRDS("../Appendix_AgeAcceleration_RNAseq/www/r1_rnaseq.RDS")
X <- lm( t(apxdata$rnaseq %>% na.omit) ~ ., 
  data = apxdata$modelData %>% select(-Age))$residuals %>% t

pd <- computeTrends(X, apxafit,
                    colP  = "P.DiagnosisControl.Age",
                    colFC = "C.DiagnosisControl.Age",
                    age   = apxdata$modelData$Age,
                    group = apxdata$modelData$Diagnosis) %>%
  .[, Type := "Appendix"] %>% 
  .[, Group := factor(Group, levels = c("Control", "PD/LBD"))]


textLayer <-
            pd[, cor.test(Age, Trend) %>% broom::tidy(), list(Group, Type, Dir)] %>%
            select(Group, Type, Dir, estimate, p.value) %>%
            setDT() %>%
            .[, label := 
                  sprintf("r = $%s$, p %s %s", 
                          format(estimate, digits=1), 
                          ifelse(p.value == 0, "<", "="),
                          texP(p.value)
                         )
            ] %>% 
            .[, Age := 20] %>% 
            .[, Trend := ifelse(Group == "Control", 0.2, 0.17)]



pB <- 
  pd %>%
  ggplot(., aes(Age, Trend, color = Group)) + 
    geom_point(alpha = 0.3, aes(shape = Group, fill = Group)) + 
    geom_smooth(method = 'lm', se = FALSE) + 
    geom_text(data = textLayer, aes(label = TeX(label, output="character")), 
      parse = TRUE, 
      hjust=0) +
    ylab("Mean normalized methylation") + 
    facet_grid(Type ~ Dir, scales = "free") + 
    scale_shape_manual(values = c(16, 17)) +
    scale_color_manual(values = c("black", "#e31a1c")) +
    scale_fill_manual(values = c("black", "#e31a1c")) +
    # scale_y_continuous(limits = c(-0.8, 1)) + 
    theme_bw(base_size = 14)


# Age deceleration
predicted <- readRDS("../Appendix_AgeAcceleration_RNAseq/r1_ageAcceleration.RDS")
pC <- predicted[, 
  list(
    `P-A` = P-A,
    Group = Diagnosis
  )
] %>%
.[, Group := factor(Group, 
                levels = c("Control", "PD/LBD"), 
                labels = c("Control", "PD"))] %>%
  ggplot(., aes(Group, `P-A`, color = Group)) + 
    geom_boxplot(outlier.size = 0) + 
    geom_jitter(size = 0.5, shape = 1, alpha = 0.3) + 
    stat_compare_means(method = "t.test", 
      aes(label = paste0("p = ", ..p.format..))) +
    xlab("Group") + 
    ylab("Predicted - Actual Age, years") + 
    theme_bw(base_size = 14) +
    scale_color_manual(values = c("black", "#e31a1c")) +
    theme(
      legend.position="top"
      )
```

```{r, include=TRUE, fig.cap="**Figure S10. Aging patterns of ALP gene transcripts in the healthy and PD appendix.**"}

legendGroup <- get_legend(pC)
ggdraw() +
  # draw_plot(
  #   pA,
  #   x = 0, y = 0, width = 0.25, height = 1) +
  draw_plot(pB + theme(legend.position = "none"),
    x = 0.0, y = 0, width = 0.8, height = 1) +
  draw_plot(pC + guides(color=FALSE, shape=FALSE) + 
    theme(legend.position = "top"),
    x = 0.8, y = 0, width = 0.2, height = 1) +
  draw_plot_label(
    label = LETTERS[1:2],
    size = 10,
    x = c(0, 0.8),
    y = c(1, 1))

```


