---
title: "Results Appendix"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: 
      collapsed: false
    includes:
      in_header: ../include/in_header.html
      before_body: ../include/before_body.html
      after_body: ../include/after_body.html
runtime: shiny_prerendered
---
<!-- # (c)  Juozas GordeviÄius -->

```{r, include=FALSE, echo=FALSE}
require(weights)
require(doParallel)
require(TxDb.Hsapiens.UCSC.hg19.knownGene)
require(org.Hs.eg.db)
require(cowplot)
source("../code/common.R")
source("../code/pformat.R")
source("../code/genomicEnrichment.R")
annotations <- cache(foo=loadHumanAnnotations,
                     fname = "annotations.RDS")
pathways <- readRDS("../Discover_Pathways/humanPathways.RDS")
```

# Appendix PD vs Controls, Padlocks

```{r}
apx <- fread("../Appendix_PDvsControls_Padlock/www/m6_Appendix_PDvsControls_Padlock.csv")
apd <- readRDS("../Appendix_PDvsControls_Padlock/www/m6_methylation.RDS")
```


```{r, include=TRUE, results='asis'}
n <- nrow(apx)
glue("- Number of loci {n}")
t <- apd$key[, table(Diagnosis)]
glue("- Number of PD cases: {t[2]} and controls: {t[1]}")
n <- apx[, sum(adj.P.Val < 0.05, na.rm=TRUE)]
glue("- Number of significant loci {n}")
n <- apx[adj.P.Val < 0.05, length(unique(Gene))]
glue("- Number of genes affected {n}")
```

Dominant direction of methylation:

```{r, include=TRUE}
t <-
	apx[, list(
						Significant = ifelse(adj.P.Val < 0.05, "Significant", "NS"), 
					 	Direction = ifelse(logFC < 0, "Hypo-M", "Hyper-M")
					 )] %>%
	.[, table(Significant, Direction)] %>%
	.[c("NS", "Significant"), c("Hypo-M", "Hyper-M")]

# Show the table
knitr::kable(t) %>%
kable_styling(bootstrap_options = c("hover"),
							full_width = FALSE)

# Show fisher test result
t %>% fisher.test %>% 
broom::tidy() %>%
knitr::kable() %>%
kable_styling(bootstrap_options = c("hover"))
```


```{r}
url <- "../Appendix_PDvsControls_Padlock/www/m6_Appendix_PDvsControls_Padlock.csv"
stopifnot(file.exists(url))
```

- **[Additional File 2](`r url`) DNA methylation changes in the appendix of PD patients relative to controls.**


## Enrichment of genomic elements

```{r}
foo <- function(analysis, prefix, annotations, t) {
  fits <- fread(glue("../{analysis}/www/{prefix}_{analysis}.csv"))
  dt1 <- getAnnotatrEnrichments(fits, annotations, qThresh = t)
  dt2 <- getAppendixChipEnrichments(fits, qThresh = t)
  dt3 <- getEpiRoadmapEnrichment(fits, qThresh = t)
  dt4 <- getPsychEncodeEnrichments(fits, qThresh = t)
  return(list(dt1, dt2, dt3, dt4))
}
analysis <- "Appendix_PDvsControls_Padlock"
prefix <- "m6"
t <- 0.05
# touch()
dtlist <- cache(
  foo = foo, fname = glue("enrichment_{analysis}.RDS"), verbose = TRUE,
  analysis = analysis, prefix= prefix, annotations = annotations,
  t = t
  )
```

General annotation of genomic elements:

```{r, include=TRUE}
dtlist[[1]] %>%
.[Type == "All"] %>%
.[, `P < 0.05` := texPstars(`P < 0.05`)] %>%
select(estimate=OR, 
       p.value=P, 
       conf.low=Lo, 
       conf.high=Hi, 
       category=Category, 
       p_stars = `P < 0.05`) %>%
kable() %>%
kable_styling(bootstrap_options = c("hover"))
```

Genomic elements obtained from ChIP-seq:

```{r, include=TRUE}
dtlist[[2]] %>%
.[, `P < 0.05` := texPstars(`P < 0.05`)] %>%
select(estimate=OR, 
       p.value=P, 
       conf.low=Lo, 
       conf.high=Hi,
       type=Type, 
       category=Category, 
       p_stars = `P < 0.05`) %>%
kable() %>%
kable_styling(bootstrap_options = c("hover"))
```


```{r}
myplots <- plotAnnotatrEnrichment(dtlist[[1]])
myplots[[1]] <- myplots[[1]] + facet_grid(Type~.)
myplots[[2]] <- myplots[[2]] + facet_grid(Type~.)
myplots[[3]] <- plotAppendixChipEnrichment(dtlist[[2]], wrap = "vertical")
```

```{r, include=TRUE, fig.cap="**Figure S2. Genomic elements enriched with differentially methylated cytosines in the PD appendix.**"}
ggarrange(myplots[[1]], myplots[[2]], myplots[[3]],
  ncol = 3, nrow = 1, labels = LETTERS[1:3],
  widths = c(4, 7, 3),
  common.legend = TRUE)
```




# Appendix PD vs Controls, RNAseq

```{r}
rna <- fread("../Appendix_PDvsControls_RNAseq/www/r1_Appendix_PDvsControls_RNAseq.csv")
apd <- readRDS("../Appendix_PDvsControls_RNAseq/www/r1_rnaseq.RDS")
```


```{r, include=TRUE, results='asis'}
n <- nrow(rna)
glue("- Number of transcripts {n}")
t <- apd$key[, table(Diagnosis)]
glue("- Number of PD cases: {t[2]} and controls: {t[1]}")
n <- rna[, sum(adj.P.Val < 0.05, na.rm=TRUE)]
glue("- Number of significant transcripts {n}")
```


Dominant direction of transcription change:

```{r, include=TRUE}
t <-
	rna[, list(
						Significant = ifelse(adj.P.Val < 0.05, "Significant", "NS"), 
					 	Direction = ifelse(logFC < 0, "Down", "Up")
					 )] %>%
	.[, table(Significant, Direction)] %>%
	.[c("NS", "Significant"), c("Down", "Up")]

# Show the table
knitr::kable(t) %>%
kable_styling(bootstrap_options = c("hover"),
							full_width = FALSE)

# Show fisher test result
t %>% fisher.test %>% 
broom::tidy() %>%
knitr::kable() %>%
kable_styling(bootstrap_options = c("hover"))
```


Dominant downregulation among the ALP transcripts:

```{r, include=TRUE}
alpGenes <- apx[, unique(Gene)]
t <-
	rna[, list(
						Significant = ifelse(adj.P.Val < 0.05 & logFC < 0, "Down-regulated", "NS"), 
					 	Gene = ifelse(SYMBOL %in% alpGenes, "ALP", "Other")
					 )] %>%
	.[, table(Significant, Gene)] %>%
	.[c("NS", "Down-regulated"), c("Other", "ALP")]

# Show the table
knitr::kable(t) %>%
kable_styling(bootstrap_options = c("hover"),
							full_width = FALSE)

# Show fisher test result
t %>% fisher.test %>% 
broom::tidy() %>%
knitr::kable() %>%
kable_styling(bootstrap_options = c("hover"))
```


```{r}
url <- "../Appendix_PDvsControls_RNAseq/www/r1_Appendix_PDvsControls_RNAseq.csv"
stopifnot(file.exists(url))
```

- **[Additional File 3](`r url`) Differentially expressed ALP genes in the appendix of PD patients relative to controls.**


# Enrichment of ALP pathways

## In appendix epigenome

```{r}
pd <- foreach(p = pathways, .combine = rbind) %do% {
  genes <- p$genes
  res1 <- apx[, table( adj.P.Val < 0.05, Gene %in% genes)] %>% 
          fisher.test %>%
          broom::tidy() %>%
          setDT %>%
          .[, Term := p$term] %>%
          .[, Name := p$name] %>%
          .[, Type := "Target"] %>% 
          .[, Dir := "All"]
  res2 <- apx[, table( adj.P.Val < 0.05 & logFC > 0, Gene %in% genes)] %>% 
          fisher.test %>%
          broom::tidy() %>%
          setDT %>%
          .[, Term := p$term] %>%
          .[, Name := p$name] %>%
          .[, Type := "Target"] %>% 
          .[, Dir := "Hyper-"]
  res3 <- apx[, table( adj.P.Val < 0.05 & logFC < 0, Gene %in% genes)] %>% 
          fisher.test %>%
          broom::tidy() %>%
          setDT %>%
          .[, Term := p$term] %>%
          .[, Name := p$name] %>%
          .[, Type := "Target"] %>% 
          .[, Dir := "Hypo-"]
  rbindlist(list(res1, res2, res3))
}
if (!dir.exists("www")) dir.create("www")
fwrite(pd, file = "www/enrichment_Pathways_DMCs.csv")
```

```{r, include = TRUE}
pd %>%
.[, `P < 0.05` := gtools::stars.pval(p.value)] %>% 
.[, list(estimate, p.value, conf.low, conf.high,
				 term = sprintf("%s (%s)", Name, Term),
				 direction = Dir, 
				 p_stars = texPstars(`P < 0.05`))] %>%
kable() %>%
kable_styling(bootstrap_options = c("hover"))
```

## In appendix transcriptome

```{r}
pd <- foreach(p = pathways, .combine = rbind) %do% {
  genes <- p$genes
  res1 <- rna[, table( P.Value < 0.05, SYMBOL %in% genes)] %>% 
          fisher.test %>%
          broom::tidy() %>%
          setDT %>%
          .[, Term := p$term] %>%
          .[, Name := p$name] %>%
          .[, Type := "RNA"] %>% 
          .[, Dir := "All"]
  res2 <- rna[, table( P.Value < 0.05 & logFC > 0, SYMBOL %in% genes)] %>% 
          fisher.test %>%
          broom::tidy() %>%
          setDT %>%
          .[, Term := p$term] %>%
          .[, Name := p$name] %>%
          .[, Type := "RNA"] %>% 
          .[, Dir := "Hyper-"]
  res3 <- rna[, table( P.Value < 0.05 & logFC < 0, SYMBOL %in% genes)] %>% 
          fisher.test %>%
          broom::tidy() %>%
          setDT %>%
          .[, Term := p$term] %>%
          .[, Name := p$name] %>%
          .[, Type := "RNA"] %>% 
          .[, Dir := "Hypo-"]
  rbindlist(list(res1, res2, res3))
}
if (!dir.exists("www")) dir.create("www")
fwrite(pd, file = "www/enrichment_Pathways_DEGs.csv")
```

```{r, include = TRUE}
pd %>%
.[, `P < 0.05` := gtools::stars.pval(p.value)] %>% 
.[, list(estimate, p.value, conf.low, conf.high,
				 term = sprintf("%s (%s)", Name, Term),
				 direction = Dir, 
				 p_stars = texPstars(`P < 0.05`))] %>%
kable() %>%
kable_styling(bootstrap_options = c("hover"))
```

<!--  -->
```{r, include=TRUE, fig.cap="**Figure S3. ALP pathway changes in DNA methylation and gene expression in the PD appendix.**"}
pd <- rbind(
  fread("www/enrichment_Pathways_DEGs.csv"), 
  fread("www/enrichment_Pathways_DMCs.csv")[, Type := "Methylation"])

p1 <- 
	pd[Type == "Methylation"] %>%
	ggplot(aes(Name, log(estimate), color = p.value < 0.05)) + 
	  geom_point(size = 4) + 
	  geom_errorbar(aes(ymin = log(conf.low), ymax = log(conf.high)), size = 0.25, width = 0.25, color = "black") + 
	  geom_hline(yintercept = 0) + 
	  ylab("OR, log") +
	  facet_grid(Type ~ Dir) + 
	  scale_color_brewer("P < 0.05", palette="Dark2") + 
	  theme_bw(base_size = 14) + 
	  theme(axis.title.x = element_blank(),
	        axis.text.x = element_blank(),
	        legend.position = "top")

levels <- c("All", "Hypo-", "Hyper-")
labels <- c("All", "Down-regulation", "Up-regulation")

p2 <- 
	pd[Type == "RNA"] %>% 
	.[, Dir := factor(Dir, levels = levels, labels = labels)] %>%
	ggplot(aes(Name, log(estimate), color = p.value < 0.05)) + 
	  geom_point(size = 4) + 
	  geom_errorbar(aes(ymin = log(conf.low), ymax = log(conf.high)), size = 0.25, width = 0.25, color = "black") + 
	  geom_hline(yintercept = 0) + 
	  ylab("OR, log") +
	  facet_grid(Type ~ Dir) + 
	  scale_color_brewer("P < 0.05", palette="Dark2") + 
	  theme_bw(base_size = 14) + 
	  theme(axis.title.x = element_blank(),
	        axis.text.x = element_text(angle = 60, hjust = 1),
	        legend.position = "none")


ggdraw() + 
	draw_plot(p1, x=0, y=0.6, width = 1, height = 0.4) +
	draw_plot(p2, x=0.015, y=0, width = 1-0.015, height = 0.6)
```

# Enrichment near lysosomal genes


```{r}
annotateTargetGenes <- function(fromFitFile) {
	# The full list of targeted genes
	targetGenes <- 
		fread(fromFitFile) %>%
		.[, unique(Gene)]

	g = genes(TxDb.Hsapiens.UCSC.hg19.knownGene)
	xx <- as.list(org.Hs.egSYMBOL2EG)

	geneCoordinates <-
		g %>% 
		as.data.table %>%
		.[, list(Chr=seqnames, Start=start, End=end, Strand=strand, ID=gene_id)] %>%
		setkey(ID)

	id2symbol <- 
		lapply(names(xx), function(gene) { data.table(Symbol=gene, ID=xx[[gene]])}) %>% 
		rbindlist %>%
		setkey(ID)

	targetCoordinates <- 
		merge(geneCoordinates, id2symbol) %>%
		.[Symbol %in% targetGenes]

	targetCoordinates[, Type := "Other targeted ALP genes"]
	targetCoordinates[Symbol %in% pathways[[3]]$genes,
											Type := "Lysosome genes"]

	targetCoordinates[, list(Symbol, Type, Chr, Start, End, Strand)]
}

loadData <- function(dataFile, fitFile) {
	met <- readRDS(dataFile)
	data <- list()
	data$key <- met$key
	data$met <- met$methylation[, as.character(data$key$Sample_ID)]
	data$genome_met <- met$genome
	data$fit <- fread(fitFile)
	data
}

computeEnrichment <- function(data, targetGenes) {
	dt <- merge(data$fit[, list(Gene, SNP, P.Value, adj.P.Val)], 
	            targetGenes[, list(Symbol, Type, Strand, Chr, Start, End)], 
	            by.x = "Gene", 
	            by.y = "Symbol", 
	            all.x = TRUE)
	dt <- dt %>% na.omit

	dt[Strand == "-", GeneStart := End]
	dt[Strand == "+", GeneStart := Start]
	dt[, Dist := abs(SNP - GeneStart) ]
	dt[, Bracket := Dist %/% 20000]
	dt[Dist > 600000, Bracket := NA]
	dt[, table(Bracket)]
	brackets <- dt[, unique(Bracket)]

	# Compute for subsets
	pd <-
		foreach( target = dt[, unique(Type)]) %do% {
			foreach( bracket = brackets) %do% {
				tryCatch({
					dt[Type == target, table(Bracket == bracket, adj.P.Val < 0.05)] %>%
					fisher.test %>%
					broom::tidy() %>%
					setDT %>%
					.[, Bracket := bracket]		
				}, error = function(e) {
					data.table(Bracket = bracket, p.value = NA)
				})
			} %>%
			rbindlist(fill=TRUE) %>%
			.[, Target := target]
		} %>%
		rbindlist(fill=TRUE)
	
	# Compute for all
	pd2 <-
		foreach( bracket = brackets) %do% {
			tryCatch({
				dt[, table(Bracket == bracket, adj.P.Val < 0.05)] %>%
				fisher.test %>%
				broom::tidy() %>%
				setDT %>%
				.[, Bracket := bracket]		
			}, error = function(e) {
				data.table(Bracket = bracket, p.value = NA)
			})
		} %>%
		rbindlist(fill=TRUE) %>% 
		.[, Target := "All"]
	pd <- rbind(pd, pd2)
	pd[, Q := p.adjust(p.value, "fdr")]
	pd[Q < 0.05]
	return(pd)
}

plotEnrichment <- function(pd) {
	pd <- pd[Bracket <= 13]
	labels <- seq(20, 600, 20)
	breaks <- 0:(length(labels)-1)
	labels <- glue("[{labels-20}, {labels})")

	pd %>%
	.[, Q := p.adjust(p.value, "fdr")] %>%
	ggplot(aes(Bracket, log(estimate), group=Target)) + 
	geom_errorbar(
								aes(ymin=log(conf.low), 
								    ymax=log(conf.high)),
								position=position_dodge(width=0.5),
								width=0.2, 
								color="grey70") + 
	geom_point(aes(color=Q<0.05), size=4, 
	           position=position_dodge(width=0.5)) + 
	geom_line(linetype=2, color="grey70",
	          position=position_dodge(width=0.5)) + 
	geom_hline(yintercept=0, color="grey") + 
	facet_wrap(~Target) + 
	scale_x_continuous("Absolute distance to gene start, kbp",
											breaks = breaks, 
											labels = labels) + 
	scale_color_brewer("FDR q < 0.05", 
	                   palette="Dark2", 
	                   direction=1 ) + 
	ylab("Odds ratio, log") +
	theme_bw(base_size=14) + 
	theme(legend.position = "top",
	      panel.grid.major = element_blank(),
	      panel.grid.minor = element_blank(),
				axis.text.x = element_text(angle=45, hjust=1))
}
```

```{r}
data <- loadData(
	"../Appendix_PDvsControls_Padlock/www/m6_methylation.RDS",
	"../Appendix_PDvsControls_Padlock/www/m6_Appendix_PDvsControls_Padlock.csv"
)
# touch()
targetGenes <- 
	cache(
		foo   = annotateTargetGenes, 
		fname = "targetsAppendix.RDS",
		fromFitFile = "../Appendix_PDvsControls_Padlock/www/m6_Appendix_PDvsControls_Padlock.csv"
	)
pd <- computeEnrichment(data, targetGenes)
```

```{r, include=TRUE}
pd[Bracket <= 9][, Q := p.adjust(p.value, "fdr")][Q < 0.2] %>%
select(-method, -alternative) %>%
kable() %>%
kable_styling(bootstrap_options="hover")
```


```{r, include=TRUE, fig.cap="**Figure S4. Proximity of DNA methylation changes to ALP genes.**"}
plotEnrichment(pd)
```



# Correlation of fold changes


```{r}
dt <- 
	merge(apx[, list(Gene, ID, Str, logFC, P.Value, adj.P.Val)],
			rna[, list(Gene = SYMBOL, Str, logFC, P.Value, adj.P.Val)],
			by = "Gene")

dt2 <- 
	dt %>% 
	.[Str.x == Str.y] %>% 
	.[, list(weighted.mean(logFC.x, -log(adj.P.Val.x), na.rm=TRUE), 
	          logFC.y[1], P.Value.y[1]), Gene]

```

N = `r nrow(dt2)` genes.

```{r, include=TRUE}
dt2[, wtd.cor(V1, V2, weight=-log(V3))] %>%
kable() %>%
kable_styling(bootstrap_options="hover")
```

